
# Planning and preparation

The chapter presents recommendations for survey designers, survey managers and trainers on how to plan and prepare an effective and efficient fieldwork training. Fieldwork training should be planned well in advance and must be well prepared. They usually consist of a main interviewer training and separate training for supervisors, data monitors or other fieldworker roles.    

Survey designers, please take note of the sub-chapters [training size](#training-size), [trainers](#trainers) and [Timing & duration](#trainingduration), as they are budget and timeline relevant. For survey managers and trainers all sub-chapters are relevant to plan and prepare a solid fieldworker training.

## Training size {#trainingsize}

LSMS strongly recommends to train more trainees than are ultimately needed as fieldworkers, i.e. as interviewers, supervisors, data monitors or other roles. This allows for a [competency-based selection](#assess-select) of fieldworkers at the end of the training. Training more people than needed also compensates any attrition of trainees during the training and builds a reserve of already trained persons that can be brought up to speed relatively quickly during fieldwork, should fieldworkers need to be replaced.

Train __at least 40% more fieldworkers__ than are required if you are working with unknown trainees, a new implementing partner or in a new context. For example, train at least 56 people if you are aiming for 40 fieldworkers. Even if some drop out before the end of the training, this still allows you to not select some of the trainees that are not up to the task. You can reduce the excess to 20% if you are working with experienced teams who you have worked with before and if you are expecting few trainees to drop out during the training. Training less than 20% additional people is not recommended as it prohibits a meaningful fieldworker selection process. 

Often, some trainees will drop out during the training, particularly after day 1 or 2 once there is a clearer understanding of the required effort. Mitigate early trainee attrition by keeping on standby some of the applicants that have not been invited to the training, so they can quickly replace any trainees that have not come back. You might have to bring them up to speed in extra hours and ask them to independently learn already covered material to catch up. It is not advised to bring in additional trainees after day 3, as they will have fallen behind too far. 

:::: {.warning data-latex=""}
In some surveys, there have been walk-outs of trainees, often towards the end of the training, in which trainees collectively refuse to continue the training or fieldwork unless their demands are met. Usually, there is no time or budget to repeat the training with other trainees, leaving survey management in a very weak bargaining position and putting the survey itself at risk. Avoid walk-outs by setting fair [fieldworker terms and conditions](Fieldworker terms), making them clear to trainees prior to the training, and asking them to agree by signing a copy of the terms and conditions. 
::::

In many circumstances it is beneficial to select supervisors, data monitors and any other fieldwork roles from the pool of trainees based on their performance and capabilities observed during the training (see [Assessment & Selection](#assess-select)). If any of the fieldwork roles are predetermined, the designated persons should participate in the fieldwork training without being counted towards the trainees excess (and demonstrate full understanding of the trained material).  

Keep the number of trainees as small as possible, __ideally below 50 trainees__, which corresponds to around 36 fieldworkers. Learning outcomes quickly decrease with increasing class size, as each trainee receives less supervision and personal feedback and as it is more difficult for trainers to keep track of trainees' performance. If possible, adjust the fieldwork plan, so that fewer fieldworkers work over a longer period of time and fewer fieldworkers need to be trained. See Chapter [Fieldworker Recruitment](#fieldworker-Recruitment) for other benefits of having a smaller field team.  

If it is unavoidable to train more than 50 trainees (e.g. surveys in large country or with large sample) or a centralized training is not possible (e.g. different languages or geographically dispersed sample), it is best or can be imperative to split fieldworker training into separate rooms or even locations. In those instances it is paramount to __standardize the training__ between different rooms/locations as much as possible in order to minimize the "[training effect](#impacterror)". 

This requires the training to be largely guided by standardized and extremely well-prepared training materials, and effective communication between the training rooms/locations, so that all trainees benefit from relevant feedback, comments, questions, etc. raised in one room/location. on top of that, there must be sufficient capable trainers who have received a comprehensive *Training of Trainers* (ToT). the ToT can be a dedicated exercise or integrated as an explicit outcome of the pilot training. Several scenarios are possible if a large number of fieldworkers need to be trained:

1. __Break-out sessions__. This is the first best option to train class sizes between 50-100 trainees if they can be trained centrally. Split the class into groups of manageable size around 15-20 trainees and assign dedicated trainers to each group. Train all modules that are sensitive to class size in the groups, including the questionnaire reading and practicing. Trainers must keep notes of any additional clarifications they provided, questions they received or feedback they gave. Hold daily plenary sessions with the entire class in which feedback is given based on the notes from individual groups. Modules that are less size-sensitive, such as the opening module or fieldwork logistics can be trained in the full class. 

1. __Parallel training__. If joint plenary session become unfeasible due to larger class size, the group of trainees must be split and trained either in parallel or consecutive training. If possible, try to avoid this by adjusting the fieldwork plan. In parallel training, groups are independently trained at the same location. Each group should have a "resident" trainer, who is responsible for the training in this group. Other trainers can rotate between groups, for example those responsible for training a certain module or task, bringing some consistency to the training. Due to groups and trainers not overlapping, it is harder to harmonize training between groups. It is therefore important that trainers debrief during the breaks and at the end of the day. 

1. __Consecutive training__. Different groups of interviewers are trained one after the other. A core set of trainers repeats the same training for each group, providing consistency between the groups. Since each group should start fieldwork immediately after their training, this approach only works if a staggered field start is possible (e.g. if field work is sufficiently long). Additional trainers should attend the training for each group, so that they can supervise the field start of one group while the core trainers move on to train the next one.    

1. __Parallel locations__. It is not recommended to simultaneously train in multiple locations, as it is difficult to harmonize fieldwork training between different locations. If unavoidable, it is essential for trainers to keep good notes, frequently communicate and debrief with trainers from other locations. 


## Trainers 

The success of a fieldwork training hinges upon having enough and competent trainers.

Conducting a fieldwork training is a substantial amount of work! Doing it well even more so. [Prepare the training well](#trainingpreparation) to avoid significant reductions in quality or trainers getting overwhelmed. Make provisions for enough trainers so they can conduct the training well. For every room, there needs to be __a minimum of two active trainers__ at any point in time, one to lead the training and another one to support the class or do any of the supporting activities that need to happen in parallel, such as updating manuals, compiling feedback, creating practicing scenarios, etc. Larger classes require more trainers as supervision and supporting activities increase. Ideally, there should be one active trainer for every 15-20 trainees.

Trainers must have availability to actively engage in the training most of the time. Avoid the situation where trainers are physically present, but either are in meetings, stare at their laptops or do other tasks unrelated to the survey. Cater for more trainers if trainers have ongoing other responsibilities outside of the training (which is often the case).

All trainers must have:

- A good understanding about the survey in general 
- A detailed understanding of the questionnaires, definitions, protocols, manual, CAPI, etc. Ideally, they have been involved in the survey design process, participated in the pre-test or pilot and will be involved in managing and monitoring fieldwork. 
- A general understanding of the survey subject, e.g. the regional agricultural systems for an agricultural survey. 
- Good facilitation skills and the ability to lead and manage all components of the training. 

Collectively, the trainers of one room should also have: 

- A good understanding of the indicators construction, the intended data use or analysis.  
- Subject and context expertise, e.g. know in detail the local circumstances of child education, nutrition and health for a survey on early child development. 
- Experience in implementing the survey type and implementing surveys in the context 
- Experience in conducting fieldworker training 
- Speak the trainees' language. Do not train in a language the trainees are not fluent in. If some trainers (usually the survey designers or analysts) do not speak the training language, use an interpreter or ask other trainers to translate for them. 

Particularly in surveys with limited pre-testing and piloting, often mismatches between the survey design and field reality are uncovered during the training and adjustments need to be made quickly. For example, a locally practiced exchange labor system might not be picked up accurately in the questionnaire. Trainers need the above expertise and experience to be able to identify those issues and take decisions that are analytically correct, in-line with the survey design and make sense in the local context. 

There is a tendency to send junior analysts or PhD students without sufficient survey or context experience to conduct the interviewer training. This can work, as long as other experienced trainers are present, but quickly leads to quality issues if they are the main training lead.  

Similarly, handing over a sample and questionnaire to a survey firm and letting them implement the training without monitoring them closely is usually a bad idea. Firms commonly lack the analytical background, do not know about the details of the survey design, are differently motivated, and might not have the capacity to conduct the training to the required standard. If working with a new firm, it is recommended to send your own trainers to be present at the training, support if necessary and ensure the training meets the quality standards described here.

Some surveys advocate for subject experts to facilitate some of the modules. Unfortunately, they often have their own agenda and priorities and come with a varying capacity to conduct effective fieldworker training, making it sometimes a waste of time (e.g. long, theoretical presentations on details that are largely irrelevant for interviewers). If external subject experts will lead parts of your training, make sure they fully understand the relevant parts of the questionnaire and what knowledge they need to transmit to trainees. Prior to the training, discuss the scope of their engagement and review their contributions, such as presentations. Subject experts can also be given a backstopping role to answer any questions on the subject and contribute to relevvant disucssions.

## Content

To equip trainees with all the knowledge and skills necessary to do a successful job as fieldworker, training must go beyond a simple discussion of the questionnaire. It should cover the following components, see details on how to implement them in the [next chapter](#conducting-the-training).

- __CAPI use__. Interviewers must understand and be able to use all functionality of the CAPI software they need to use in the field, including handling the CAPI questionnaire, creating and sending interview files, etc. it is best to keep theory to the limit and introduce functionality throughout the training as needed, so trainees can practice it immediately.

- __Questionnaire content__. All questionnaires must be covered in detail, including question text, answer types and options, how to administer questions and sections, the routing/skipping behavior, and the underlying definitions and concepts. 

- __Practicing__. Repeated practicing, in groups, with respondents on-site and in the field is crucial for trainees to fully internalize the trained material, learn how to put it into practice and to get exposure to real respondents and scenarios during the training.
 
- __Interviewing techniques__. Trainees must learn and practice a range of skills that are critical to collect good data, such as how to introduce yourself, the survey and convince respondents to participate, how to probe, inquire, and get clarifications, how to control the interview.

- __Special tasks__. Many surveys include exercises such as capacity tests, anthropometrics, plot measurement or behavioral games that require strict adherence to protocols for comparable measurements. Trainees must practice until they implement them correctly and until behavior across the entire field team has been homogenized.

- __Pre-interview skills__. Interviewers need to learn the format in which they will receive their assigned units (e.g. households), any selection protocols they might have to follow, how to identify and locate units, when to revisit, replacement protocols, how to manage their workload, etc. 

- __Post-interview skills__. It is important for interviewers to know correctly perform a range of activities after an interview, including checking for completion, addressing inconsistencies, leaving comments, submitting files, receiving feedback, respond to inquiries on submitted interviews, making corrections, asking questions, etc.

- __Final field test__. At the end of the training, a final field test should be conducted to identify and address any outstanding issues with questionnaires or field procedures, expose interviewers to real field conditions and determine if they are ready for field work. 
__Supervisor training__: Supervisors must learn how to manage field teams, manage the relationship to the communities and respondents, and the tasks they are responsible for in the survey, such as handling money, making logistical arrangements, checking completions, communication to survey management or data monitors, etc. 

__Data monitor training__: Data monitors need to learn all the tools and background information necessary to be effective in their roles, including how to review interviews, conduct re-interviews review audio, how to interpret summarized information, how to provide effective feedback to fieldworkers, how to compile feedback for survey management and archiving. 

## Timing & duration {#trainingduration}

Fieldwork training should be held __just prior to the start of fieldwork__, with a maximum of 1-3 days between training and fieldwork to allow for administrative matters, a rest day travel to the field. An immediate fieldstart with strong supervision and feedback in the first few days is the best way to ensure the survey is being implemented as trained and that fieldworker behavior is streamlined. 

The longer the break, the higher is the risk of interviewers forgetting details, making mistakes and developing idiosyncratic behavior. If fieldwork is delayed by more than a few days, conduct a short __refresher training__ just prior to the field start to revisit and practice key points of the trained material. For longer delays, a longer refresher training will be necessary. For breaks beyond a few weeks, often a complete retraining is the only solution, also because there tend to be (some) new trainees.

__Schedule enough time for fieldwork training__. The optimal duration depends on a range of factors, including the nature of the questionnaire, tasks required by fieldworkers, level of trainees, etc. As a general rule of thumb, for a LSMS-style survey with a 2-2.5 hours questionnaire allow around 2.5 weeks (15 working days) for the main interviewer training, including practicing and final field tests. Training duration can vary to some degree with survey factors, but keep in mind that many components such as opening session, CAPI use, interviewing skills, final field tests need be trained independently, so even for a survey with a relatively short 45-minutes questionnaire, at least 1 week (6 working days) of interviewing training should be scheduled.   

To some readers, this might sound excessively long. Many survey firms, especially with a background in opinion surveys will suggest much shorter training duration. However, in socio economic surveys, due to the substantial amount of [content](#content) that fieldworkers need to learn and be able to apply, it is usually impossible to train the field teams to an acceptable level in a shorter amount of time. Do not save on training duration, unless you are very confident that all trainees will have fully internalized all material and will be able to perform well as field workers!  

Be prepared for possible extensions. In many surveys, trainees are not ready to go to the field at the end of the training and the training needs to be extended for a few days. If extensions have not been budgeted for, this can cause quality issues elsewhere, e.g. the survey firm having to cut corners elsewhere to compensate for increased costs. Avoid this by setting out with a realistic training schedule and putting contingencies for potential extensions into the budget and timelines.

Avoid fieldworker training that last longer than weeks (18 working days). Trainees' ability to absorb and remember material and general morale decreases with training length. Also, since every day of fieldworker training is quite expensive, paying for days that are increasingly ineffective, longer trainings become increasingly inefficient and can reduce the funds available for other important survey activities. Conducting an efficient and yet effective training as described in Chapter \@ref(conducting-the-training) [Conducting the training](#conducting-the-training) should allow most surveys to be trained well in less than 3 weeks.

For surveys with a large number of questionnaires (e.g. school surveys) or special tasks such as anthropometrics or plot measurements that take a substantial amount of time to train (e.g. schedule 4-6 days for a solid anthropometrics training), often it is unrealistic for interviewers to learn all questionnaires and tasks sufficiently well in the time available. Instead, divide questionnaires and tasks between __specialized fieldworkers roles__, e.g. anthropometrics to be taken by dedicated measurers and interviews to be conducted by interviewers. This way, each fieldworker role can focus in their questionnaires/tasks and learn how to implement them well. Different fieldworker roles can be trained in parallel training sessions, which cuts overall training length and costs. 

__Supervisor training__ should be held outside of main interviewer training hours, so that supervisor (candidates) can attend the interviewer training in full. If supervisors are known prior to training, supervisor training can be held just prior to the main training for a couple of days. If supervisors are selected during the training, it needs to be held towards the end of the training in the late afternoon/evenings or during the day in modules that are not relevant for supervisors (e.g. administrative days, revision of material they demonstrated to know, etc). The duration depends on the scope of the tasks, but generally 2-3 days are sufficient. 

__Data monitor training__ is best held after the final field test, so its data can be used as training and practicing material. Data monitors should attend interviewer training and field testing to get a good understanding of the activities they will monitor later. Availability of trainers permitting, it can be held in parallel to the supervisor training, between training and field work, and during the first few days of field work. Generally, 2-3 days are sufficient for initial training and practicing and an additional 2-3 days of supervised work during the first few days of field work. 

## Schedule {#trainingschedule}

Prepare a realistic training schedule and track training progress on an on-going basis. Keep in mind the following general considerations: 

1. __Stay flexible__. Create a schedule that identifies all modules and allocates the required time. Use the schedule to track progress, ensure you are not falling behind and to plan your logistics, e.g. on what days you will need the training venue or transport. Expect timings to shift a little during the training, as some things tend to take longer or shorter than anticipated. Do not force through the original schedule, but adjust the training schedule to the updated requirements of the training. Stay flexible to allow sufficient time to address issues, retrain topics, or dive into more detail as needed. 

1. __Allocate time generously__. It is better to finish training early one day rather than keeping trainees until late, rushing content or trainees not being ready for the field at the end. If you are falling behind schedule during the training, extend the training as early as possible before logistical arrangements make this impossible. If training progresses faster than anticipated, use the time for additional practicing. Only shorten the training if you are certain that the trainees are fully ready for the field.

1. __Do not exceed 6-7 hours of training time per day__ and give frequent breaks, without the breaks taking over the day. For example, 7 hours per day can translate to a morning session from 8.30 AM - 12 PM, and an afternoon session from 1 PM - 5 PM, each with a flexible 15 minutes break. Close the main training early enough to allow for supervisor training, daily trainer debriefs and preparing for the following days. 

1. __Frequently change training modality__, to keep trainees engaged and able to absorb and solidify content. For example, regularily break long blocks of content training with  practicing or interviewing skill role games. 

1. __Schedule around low energy periods__. Schedule theory and important content in the mornings when trainees tend to be freshest. Trainee attention tends to decrease in the afternoon, and is at its lowest after coming back form the field or directly after having a heavy lunch. During those periods, schedule hands-on and interactive modules that require greater engagement from trainees, such as practicing, interviewing techniques or written tests. Avoid training after coming back from the field.    

1. __Train what is relevant__. Only train content that interviewers need to know. Allocate time on topics proportionate to the direct importance for the interviewers. For example, spent less time on the general background of the survey and more on how to introduce it properly.    

1. __Train when needed__. Train material just before it is needed and can be practiced. Otherwise, trainees often forget and the content needs to be retrained. For example, do not train the household IDs, household identification, and replacement protocols at the beginning of the training, but just prior to the final field test, once trainees already are confident with the questionnaire, have the capacity to focus on the new content and are about to use it in the field  for the first time. 

1. __Train supervisors and data monitors outside of main training hours.__ They need to attend the main fieldworker training in full to understand field procedures. Typical time slots are prior to the main training (if roles are predetermined) or towards the end of the training once roles have been selected, during the late afternoon or on administrative, travel or rest days.    

1. __Schedule time to debrief after field days__. Avoid multiple testing or practicing days back to back. Trainees can get used to their idiosyncratic, potentially undesired behavior. After a field practice or field test, allow for enough time to debrief, provide feedback and retrain topics that have not gone well yet, so trainees can try to do better on the next day. Typically, one day of in-calss training is needed between field days.

1. __Rest days__. For training lasting over one week make sure to schedule one or two rest days per week. Trainees and trainers need regular breaks for good learning outcomes.  

The training schedule in Table \@ref(tab:trainingschedule) illustrates how an effective and efficient fieldwork training can be organised for a survey with 2.5-3 hr questionnaire. The schedule splits the questionnaire into two parts (e.g. household and agricultural questionnaires). First, one part is covered and extensively practiced on-site and in the field, allowing trainees to solidify the material. The same is repeated for the second part of the questionnaire, before field practices and final field test at the end bring it all together. 

The schedule follows the above general principles and includes daily written tests and reviews. The number in brackets reflect the indicative hours allocated for each module. Hover over each day to find out more details on the reasoning behind the sample schedule. 
You can download the sample schedule as an Excel version from [here](resources/training_schedule.xlsx). 

```{r training_schedule__alt, echo = FALSE, warning = FALSE, comment = FALSE, message = FALSE}
library(kableExtra)
options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = '')
library(readxl)

schedule <- read_xlsx("resources/training_schedule.xlsx", 
                      sheet = "new",
                      col_types = c("numeric","text","text","text"))
schedule$Main <-  cell_spec(
    schedule$Main,
    popover = spec_popover(
      content = schedule$Explanation,
      title = NULL,  # title will add a Title Panel on top
      position = 'auto')
    )
days <- rle(schedule$Day)$lengths
names(days) <- paste("Day",rle(schedule$Day)$values)

schedule[,1:3] %>%
  kable(escape = FALSE, 
        align="cll",
        caption = 'An Illustrative Training Schedule.', 
        label = "trainingschedule") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                fixed_thead = T) %>%
  column_spec(1, bold = T)

```

OLD FORMAT, WHICH ONE DO WE PREFER? (IGNORE CONTENT)

```{r training_schedule, echo = FALSE, warning = FALSE, comment = FALSE, message = FALSE}
library(kableExtra)
options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = '')
library(readxl)

schedule <- read_xlsx("resources/training_schedule.xlsx", 
                      sheet = "old",
                      col_types = c("numeric","text","text","text","text"))
schedule$Module <-  cell_spec(
    schedule$Module,
    popover = spec_popover(
      content = schedule$Explanation,
      title = NULL,  # title will add a Title Panel on top
      position = 'auto')
    )
days <- rle(schedule$Day)$lengths
names(days) <- paste("Day",rle(schedule$Day)$values)

schedule[,2:4] %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  pack_rows(index = days, label_row_css = "background-color: #666; color: #fff;")
```

new old 

```{r training_schedule_2, echo = FALSE, warning = FALSE, comment = FALSE, message = FALSE}
library(kableExtra)
options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = '')
library(readxl)

schedule <- read_xlsx("resources/training_schedule.xlsx", 
                      sheet = "new",
                      col_types = c("numeric","text","text","text"))
schedule$Module <-  cell_spec(
    schedule$Module,
    popover = spec_popover(
      content = schedule$Explanation,
      title = NULL,  # title will add a Title Panel on top
      position = 'auto')
    )
days <- rle(schedule$Day)$lengths
names(days) <- paste("Day",rle(schedule$Day)$values)

schedule[,1:3] %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(1, bold = T)
```

## Location

Fieldwork training is best held in a location that:

- __Has sufficient logistically capacity__: Is easy to reach for trainers and trainees, has a suitable training venue, sufficient accommodation and food options, good network coverage, print shops, etc. 

- __facilitates field testing__: Is close to communities that are similar to but outside of the sample. If the survey is limited to a specific region or area, it is best to conduct the fieldwork training there.

- __facilitates field start__: Allows field teams to start field work shortly after the training and trainers to supervise the first few days of field work. 

Find a training venue with a big hall large enough to comfortably fit the entire team (or multiple halls if parallel sessions are held) and space to breakout into smaller teams for group exercises, such as a garden, hallways, the cantine or additional rooms. It is crucial that there is __good internet connectivity__ at the venue, either through on-site Wifi, or general network connectivity. 

The venue should be either able to accommodate trainees and cater for food, or be easily accessible and close to restaurants or eateries. It should be quiet and free from distractions. Good venues tend to be hotels or conference centers with garden or open space. Avoid universities, schools, towns halls or other places that are busy and become easily unavailable. 

Book the venue early. Check for availability, in case the training needs to be extended on short notice. 


## Preparation {#trainingpreparation}

Avoid wasting precious training time by preparing everything that can be prepared prior to fieldwork training. Prior to the training, the following should be in place: 

- __Inform trainees of survey details and terms__. To minimize attrition of trainees, make sure that trainees have been informed of and agreed to the overall survey parameters and contract terms, including the selection process, their remuneration package, insurance coverage, areas to be visited, the length of field work, the transport and lodging arrangements during field work, expected effort, etc. Often, trainees drop out from the training if their expectations were wrong. Have trainees sign a copy of the details prior to the trianing to indicate agreement. 

- __Contractual matters__. All contracts, payment arrangements, insurance policies and other administrative matters should be prepared such that they do not delay training or fieldwork and that trainees are paid without delays and are covered by insurance. Avoid interrupting the training with contractual matters, e.g. by pulling out individual trainees to sign contracts. Instead, schedule dedicated admin slots prior to the training, after class or between training and fieldwork. For fieldworker selection at the end of the training, make sure that all contractual matters including insurance and payments can be addressed in the short time window between training and field start.  

- __Logistical/administrative support__. Get a dedicated person to support with all the logistical and administrative work that has to be done during the training, e.g. setting up the venue, organizing transport, contacting local leaders, printing, purchasing equipment, signing fieldworker contracts, etc. If these activities have to be done by the trainers, they become a substantial distraction and can significantly affect training quality and pace.

- __Projector with a BIG screen__. The projected content must be visible and legible to all trainees. You should be able to connect a PC and tablet/phone to the projector. Ideally, it should be possible to quickly link trainees' devices to the projector, to facilitate exercises and practicing. Test the set-up prior to the start of the training.

- __Stable internet connection for all devices__. Trainees have to sync their devices several times per day. Usually, the WiFi connection of the training venue is not sufficient to connect all devices at the same time. Often, additional mobile network routers or WIFI hot-spots are required. One option is to already provide the internet connection that is planned for field work, e.g. the dongles, mobile phone routers or sim cards.  

- __Food and refreshments for the breaks__. If food is provided by the venue, make sure it is ready in time for the break. If trainees eat out for lunch, set reasonable break times that allow trainees to finish lunch and to be back in time. Incentivize punctuality, e.g. those coming back late from the break have to sing a song.  

- __One tablet/phone per trainee__. It is crucial for all trainees to learn how to navigate the questionnaire CAPI and use the device. If you do not have enough devices (since there are more trainees than interviewers) borrow devices or ask some trainees to use their private devices. Devices must meet the minimum specifications for the CAPI software (see [here](https://docs.mysurvey.solutions/faq/what-tablets-should-i-buy-/){target="_blank"} for Survey Solutions) and have the correct CAPI version installed. Consider using an app blocker to allow using specific apps only. Add shortcut/icons on the main screen for CAPI, the softcopy of the manual, the calculator or any other relevant app. 

- __Fully set-up CAPI software__. For Survey Solutions this includes the installation of Survey Solutions on a server, interviewer accounts for each trainee, questionnaires imported to the server and assignments made. For training, often open assignments with infinite quantity (-1) are useful. Create several assignments with different scenarios if the questionnaire includes substantial pre-loading (e.g. household rosters in a panel survey). 

- __Electricity__. Avoid delays due to power issues. Provide (lots of) extension cords to charge tablets during the training and/or in the evenings. Provide power banks for field days if the device charge does not last a full day.

- __Samples of documents__. If fieldworkers need to interpret, verify or extract information from any documents or items such as immunization cards, school timetables, or anti-malaria medications, get enough copies/samples to allow for hands-on practicing. If there is variation, provide different examples, e.g. different antimalarial medication brands. 

- __Measurement equipment__. If fieldworkers need to use any special measurement equipment, such as height/length boards, GPS units, water testing kits, student testing booklets, etc., provide enough units, so that trainees can practice in a meaningful way. 

- __Field procedure documents__. Sufficient copies of any documents other than instrument and manuals that fieldworkers need to use in the field, such as introduction letters, household tracking lists, maps, cluster completion sheets, referral forms, etc. 
 
- __Finalized questionnaires__. The questionnaires should be reviewed, extensively pre-tested or piloted, translated and the translations verified. The CAPI version of the questionnaire should be extensively desk-tested and error free. Mistakes can cause significant issues or delays during the training. Provide printed copies of the paper questionnaire [if needed](#capi-use). 

- __Fieldworker manuals__. All manuals for interviewer, supervisor or other fieldworker roles have to be made available to trainees from the beginning of the training, so trainees can use them and get used to using them as a reference. [Modifications and additions](#questionnaire-content) should be made to the manuals during the training wherever it was unclear or incorrect. Either hand out a printed version of the manual at the beginning and a revised version at the end, or provide regularly updated soft copies in pdf to the trainees which they can look at on their devices (which tends to work very well). 

- __Organize in-field practicing__. Select practice areas/communities well in advance. They must not be included in the sample. Make all necessary administrative arrangements such as getting permissions, introduction to village officials, etc. If necessary, arrange respondents for field practices, so trainees can spend more time practicing than looking for respondents to participate (don't do this in areas/communities for the final field test). If you require certain types of respondents (e.g. children for athropomoterics) check with institutions where they are concentrated, e.g. child clinics, schools, etc. Organize reliable transport for the trainees to travel to the field practice. Make sure the transport waits for the team, not the other way round.

- __Organize on-site practicing__. Arrange respondents for on-site practicing to visit the training venue. Arrange for their transport and remuneration and don't forget to also provide food/refreshment for them.    

- __Facilitation material__. Microphones and speakers including batteries and fully tested, flip charts with non-permanent markers in multiple colors for drawings, name tags/stickers for all participants, printed attendance lists and feedback forms, for trainees notebooks, pens in 2 colors to allow  marking and clipboards, etc. It tends to e useful to have a printer or printing facilities in the venue to allow ad-hoc printing of e.g. exercises, tests, etc.   

- __Field work logistics and material__. All necessary logistics need to be in place to allow fieldwork to commence immediately after training, including fieldworker contracts to be signed after selection, insurance policies, payment arrangements, transport, accommodation, etc. Similarly, have all fieldwork equipment ready, such as bag packs, rain coats, ID cards, battery packs, t-shirts, respondent incentives, etc. 

